Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over \cye{convex domains and} domains with shellable triangulations

### 2024-09-27 main.tex
outlook and basic combinatorics

We gather basic notions and definitions concerning simplicial meshes.

@@

The strictly positive convex combinations of the vertices of the simplex constitute the \textit{ interior} of the simplex,
and its remaining points constitute the \textit{boundary} of the simplex.
\\

A finite family of simplices $\calT$ is a \emph{simplicial complex} or \emph{triangulation} 
if it satisfies the following conditions: 
(i) $\calT$ contains all the subsimplices of its members (ii) any non-empty intersection of two members of $\calT$ is a common subsimplex of each of them. 
% (i) every subsimplex of any simplex in $\calT$ is itself part of $\calT$, and (ii) the intersection between any two simplices $T, T' \in \calT$ is either empty or a shared subsimplex of both $T$ and $T'$. 
We say that a simplicial complex $\calT$ has \textit{dimension $n$} or is \textit{$n$-dimensional} if each of its simplices is a subset of an $n$-dimensional member of that triangulation.\footnote{Simplicial complexes that we call $n$-dimensional are called purely $n$-dimensional in the literature on polytopes~(cf.\ \cite{ziegler2012lectures}).} 
We also write $\underlying{\calT}$ for the \textit{underlying set} of the simplicial complex $\calT$, which is the union 
\begin{align*}
    \underlying{\calT} := \bigcup_{ T \in \calT } T
    .
\end{align*}
We call any set \textit{trianguable} if it is the underlying set of some triangulation. 
\\

@@

Whenever $\calT$ is a simplicial complex, 
the set of $k$-dimensional simplices in $\calT$ is denoted as $\subsimplex_{{k}}(\calT)$. 
% Additionally, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Edges(\calT) := \subsimplex_{1}(\calT)$ refer to the vertices and edges of the simplicial complex, respectively. In practice, we not always distinguish between points and singleton simplices. 
% Similarly, if $\calT$ is an $n$-dimensional simplicial complex, we write $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ for the faces (that is, codimension one members) of this triangulation.\footnote{
Similarly, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ refer to the vertices and the faces (that is, codimension one members) of this triangulation.\footnote{
% \begin{remark}
    Our use of the term \textit{face} as is common in the finite element literature~\cite{brenner2008mathematical}
    and is synonymous with \textit{facet} as used in the literature on polyhedral combinatorics.~\cite{schrijver1998theory}
    Notably, this terminology differs from the uses \textit{face} and \textit{facet} in the theory of polyhedra.~\cite{ziegler2012lectures}
% \end{remark}
}
% Given $T \in \calT$, we let $\supsimplex(\calT,T)$ be the set of simplices in $\calT$ that contain $T$,
% and we write $\supsimplex_{k}(\calT,T)$ for the set of $k$-dimensional simplices in $\supsimplex(\calT,T)$.\footnote{\mwl{MWL: It might be more economical to get rid of the supersimplex set, it is only used in this section to define patches.}}

@@

We introduce a specific notion of connectivity. 
Given an $n$-dimensional simplicial complex $\calT$, 
we call $n$-simplices $S,S' \in \calT$ \emph{strongly connected in $\calT$} if there exists a sequence $S_0=S,S_1,\dots,S'=S_m \in \calT$ such that $S_{i} \cap S_{i-1}$ is a face of both $S_{i}$ and $S_{i-1}$ for all $1 \leq i \leq m$. % TODO: Theo suggests reverse order of that sequence 
We call such a sequence a \emph{strong path} from $S$ to $S'$ in $\calT$. 
Clearly, strongly connected in $\calT$ is an equivalence relation among simplices. 
A \emph{strongly connected component} of $\calT$ is an equivalence class under this equivalence relation, 
and we call $\calT$ \emph{strongly connected} if it has only one strongly connected component. 
\\

@@

For the agenda of this section, the following combinatorial notion of triangulations will be helpful. 
Whenever $\calT$ is a strongly connected $n$-dimensional triangulation, 
a \emph{spanning tree} is an enumeration $T_{0}, \dots, T_{M}$ of all its $n$-dimensional simplices
where for each index $1 \leq m \leq M$ we have fixed index $0 \leq j(m) \leq m-1$ 
such that $T_{m}$ and $T_{j(m)}$ share a common face of dimension $n-1$.
We call $T_{j(m)}$ the \emph{predecessor} of $T_{m}$ within the spanning tree.

@@

%  Since $\calT$ is strongly connected, there exists an enumeration of $n$-simplices $T_0, T_1, \dots, T_M$ such that for any $1 \leq m \leq M$ there exists some fixed but arbitrary $0 \leq j(m) < m$ such that $F_m := T_m \cap T_{j(m)}$ is a face of dimension $n-1$. We write $\Omega_m := \cup_{j=0}^{m} T_m$. Hence $\Omega = \Omega_M$.

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:gradzwei}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a strongly connected finite triangulation $\calT$.
    Let $T_0 \in \calT$ be an $n$-simplex such that for all $n$-simplices $T \in \calT$ 
    we fix a strong path $T_0, T_1, \dots, T_m = T$.
    For any $1 \leq l \leq m$, let $F_{l} = T_{l} \cap T_{l-1}$.
    Then for any $u \in W^{1,p}(\Omega)$ 
    there exists $w \in W^{1,p}(\Omega)$ with $\nabla u = \nabla w$ 
    and such that for all $T \in \calT$:
    \begin{align*}
        \| w_{m} \|_{L^{p}(T_m)}
%         &\leq 
%         \volumeratio(\calT)^{\frac m p}
%         C_{\PF,T_{0},p} 
%         \| \nabla u \|_{L^{p}( T_{0} )}
%         +
%         \sum_{l=1}^{m} 
%         \volumeratio(\calT)^{\frac {m-l} p}
%         C_{\PF,T_{l},F_{l},p} 
%         \| \nabla u \|_{L^{p}(T_{l})} 
%         \\&\qquad 
%         +
%         \sum_{l=1}^{m} 
%         \volumeratio(\calT)^{\frac {m-l} p} \reflectionestimate(\calT)
%         C_{\PF,T_{l},F_{l},p} 
%         \| \nabla u \|_{L^{p}(T_{l-1})} 
%         \\
        &\leq 
        \volumeratio(\calT)^{\frac m p}
        C_{\PF,T_{0},p} 
        \| \nabla u \|_{L^{p}( T_{0} )}
        \\&\quad 
        +
        \sum_{l=1}^{m} 
        \volumeratio(\calT)^{\frac {m-l} p}
        C_{\PF,T_{l},F_{l},p} 
        \left( 
            \| \nabla u \|_{L^{p}(T_{l})} 
            +
            \volumeratio(\calT)^{\frac {1} p}
            \reflectionestimate(\calT)
            \| \nabla u \|_{L^{p}(T_{l-1})} 
        \right)
        .
    \end{align*}
%     \begin{align*}
%         \| w_{m} \|_{L^{p}(T_m)}
%         &\leq 
%         \sum_{l=1}^{m} 
%         \volumeratio(\calT)^{\frac{m-l}{p}}
%         \left( 1 + \volumeratio(\calT)^{\frac 1 p} \reflectionestimate(\calT) \right) 
%         C_{\PF,T_{l},F_{l},p} 
%         \| \nabla u \|_{L^{p}( T_{l} )}
%         +
%         \volumeratio(\calT)^{\frac m p}
%         C_{\PF,T_{0},p} 
%         \| \nabla u \|_{L^{p}( T_{0} )}
%         .
%     \end{align*}
%     \color{red}Precise form of the inequality yet to be discussed, though the construction principle is clear.
\end{theorem}

@@

\begin{remark}
    Not all triangu\cye{l}able sets admit a triangulation that is shellable. 
    Moreover, even if a set admits a shellable triangulation, not all of its triangulations \cye{are} be shellable. \cye{For example, if we extend the non-shellable triangulation of a tetrahedron from~\cite[Example 8.9]{ziegler1995lectures}} to a triangulation of a hypertetrahedron by suspending it from a new point $v_\star$, 
    then the resulting new triangulation is non-shellable and coincides with the patch around $v_\star$.
    This demonstrates that patches around boundary simplices are not necessarily shellable when the \cye{space} dimension \cye{$n$} is larger than three.
\end{remark}\todo{MV: Moved up a bit.}

@@

\begin{remark}
    Not all trianguable sets admit a triangulation that is also shellable. 
    Moreover, even if a set admits a shellable triangulation, not all of its triangulations might be shellable. 
    We refer to~\cite[Example 8.9]{ziegler2012lectures} for an examples of non-shellable triangulations of cubes and tetrahedra in three dimensions.

@@

>>>>

We gather here basic notions and definitions concerning simplicial meshes.

\subsection{\cye{A triangulation or a simplicial complex}}

@@

The strictly positive convex combinations of the vertices of the simplex constitute the \textit{ interior} of the simplex,
and its remaining points constitute the \textit{boundary} of the simplex.

A finite family of simplices $\calT$ is a \emph{simplicial complex} or \emph{triangulation} if it satisfies the following conditions: 
(i) $\calT$ contains all the subsimplices of its members (ii) any non-empty intersection of two members of $\calT$ is a common subsimplex of each of them. 
% (i) every subsimplex of any simplex in $\calT$ is itself part of $\calT$, and (ii) the intersection between any two simplices $T, T' \in \calT$ is either empty or a shared subsimplex of both $T$ and $T'$. 
We say that a simplicial complex $\calT$ has \textit{dimension $n$} or is \textit{$n$-dimensional} if each of its simplices is a subset of an $n$-dimensional member of that triangulation.\todo{MV: Don't we only use $n$-dimensional simplicial complexes? That would be a simplification of our presentation.}\footnote{Simplicial complexes that we call $n$-dimensional are called purely $n$-dimensional in the literature on polytopes~(cf.\ \cite{ziegler1995lectures}) \cye{and simply ``simplicial meshes'' in the finite element literature}.} 
We also write $\underlying{\calT}$ for the \textit{underlying set} of the simplicial complex $\calT$, which is \sout{defined as} the union 
\begin{align*}
    \underlying{\calT} := \bigcup_{ T \in \calT } T
    .
\end{align*}
We call any set triangulable if it is the underlying set of some triangulation. 
\\

@@

Whenever $\calT$ is a simplicial complex, 
the set of $k$-dimensional simplices in $\calT$ is denoted as $\subsimplex_{{k}}(\calT)$. 
% Additionally, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Edges(\calT) := \subsimplex_{1}(\calT)$ refer to the vertices and edges of the simplicial complex, respectively. In practice, we not always distinguish between points and singleton simplices. 
% Similarly, if $\calT$ is an $n$-dimensional simplicial complex, we write $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ for the faces (that is, codimension one members) of this triangulation.\footnote{
Similarly, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ refer to the vertices and the faces (that is, codimension one members) of this triangulation.\footnote{
% \begin{remark}
    Our use of the term \textit{face} as is common in the finite element literature~\cite{brenner2008mathematical}
    and is synonymous with \textit{facet} as used in the literature on polyhedral combinatorics~\cite{schrijver1998theory}.
    Notably, this terminology differs from the uses \textit{face} and \textit{facet} in the theory of polyhedra~\cite{ziegler2012lectures}.
% \end{remark}
}
% Given $T \in \calT$, we let $\supsimplex(\calT,T)$ be the set of simplices in $\calT$ that contain $T$,
% and we write $\supsimplex_{k}(\calT,T)$ for the set of $k$-dimensional simplices in $\supsimplex(\calT,T)$.\footnote{\mwl{MWL: It might be more economical to get rid of the supersimplex set, it is only used in this section to define patches.}}

@@

We introduce a specific notion of connectivity. 
Given an $n$-dimensional simplicial complex $\calT$, 
we call $n$-simplices $S,S' \in \calT$ \emph{face-connected in $\calT$} if there exists a sequence $S_0=S,S_1,\dots,S'=S_m \in \calT$ such that $S_{i} \cap S_{i-1}$ is a face of both $S_{i}$ and $S_{i-1}$ for all $1 \leq i \leq m$. % TODO: Theo suggests reverse order of that sequence 
We call such a sequence a \emph{strong path} from $S$ to $S'$ in $\calT$. 
Clearly, face-connected in $\calT$ is an equivalence relation among simplices. 
A \emph{face-connected component} of $\calT$ is an equivalence class under this equivalence relation, 
and we call $\calT$ \emph{face-connected} if it has only one face-connected component. 
\\

@@

For the agenda of this section, the following combinatorial notion of triangulations will be helpful. 
Whenever $\calT$ is a face-connected $n$-dimensional triangulation, 
a \emph{spanning tree} is an enumeration $T_{0}, \dots, T_{M}$ of all its $n$-dimensional simplices
where for each index $1 \leq m \leq M$ we have fixed index $0 \leq j(m) \leq m-1$ 
such that $T_{m}$ and $T_{j(m)}$ share a common face of dimension $n-1$.
We call $T_{j(m)}$ the \emph{predecessor} of $T_{m}$ within the spanning tree.

@@

%  Since $\calT$ is face-connected, there exists an enumeration of $n$-simplices $T_0, T_1, \dots, T_M$ such that for any $1 \leq m \leq M$ there exists some fixed but arbitrary $0 \leq j(m) < m$ such that $F_m := T_m \cap T_{j(m)}$ is a face of dimension $n-1$. We write $\Omega_m := \cup_{j=0}^{m} T_m$. Hence $\Omega = \Omega_M$.

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:gradzwei}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a face-connected finite triangulation $\calT$.
    Let $T_0 \in \calT$ be an $n$-simplex such that for all $n$-simplices $T \in \calT$ 
    we fix a strong path $T_0, T_1, \dots, T_m = T$.
    For any $1 \leq l \leq m$, let $F_{l} = T_{l} \cap T_{l-1}$.
    Then for any $u \in W^{1,p}(\Omega)$ 
    there exists $w \in W^{1,p}(\Omega)$ with $\nabla u = \nabla w$ 
    and such that for all $T \in \calT$:
    \begin{align*}
        \| w_{m} \|_{L^{p}(T_m)}
%         &\leq 
%         \volumeratio(\calT)^{\frac m p}
%         C_{\PF,T_{0},p} 
%         \| \nabla u \|_{L^{p}( T_{0} )}
%         +
%         \sum_{l=1}^{m} 
%         \volumeratio(\calT)^{\frac {m-l} p}
%         C_{\PF,T_{l},F_{l},p} 
%         \| \nabla u \|_{L^{p}(T_{l})} 
%         \\&\qquad 
%         +
%         \sum_{l=1}^{m} 
%         \volumeratio(\calT)^{\frac {m-l} p} \reflectionestimate(\calT)
%         C_{\PF,T_{l},F_{l},p} 
%         \| \nabla u \|_{L^{p}(T_{l-1})} 
%         \\
        &\leq 
        \volumeratio(\calT)^{\frac m p}
        C_{\PF,T_{0},p} 
        \| \nabla u \|_{L^{p}( T_{0} )}
        \\&\quad 
        +
        \sum_{l=1}^{m} 
        \volumeratio(\calT)^{\frac {m-l} p}
        C_{\PF,T_{l},F_{l},p} 
        \left( 
            \| \nabla u \|_{L^{p}(T_{l})} 
            +
            \volumeratio(\calT)^{\frac {1} p}
            \reflectionestimate(\calT)
            \| \nabla u \|_{L^{p}(T_{l-1})} 
        \right)
        .
    \end{align*}
%     \begin{align*}
%         \| w_{m} \|_{L^{p}(T_m)}
%         &\leq 
%         \sum_{l=1}^{m} 
%         \volumeratio(\calT)^{\frac{m-l}{p}}
%         \left( 1 + \volumeratio(\calT)^{\frac 1 p} \reflectionestimate(\calT) \right) 
%         C_{\PF,T_{l},F_{l},p} 
%         \| \nabla u \|_{L^{p}( T_{l} )}
%         +
%         \volumeratio(\calT)^{\frac m p}
%         C_{\PF,T_{0},p} 
%         \| \nabla u \|_{L^{p}( T_{0} )}
%         .
%     \end{align*}
%     \color{red}Precise form of the inequality yet to be discussed, though the construction principle is clear.
\end{theorem}

@@

\begin{remark}
    Not all triangulable sets admit a triangulation that is shellable. 
    Moreover, even if a set admits a shellable triangulation, not all of its triangulations \cye{are} be shellable. \cye{For example, if we extend the non-shellable triangulation of a tetrahedron from~\cite[Example 8.9]{ziegler1995lectures}} to a triangulation of a hypertetrahedron by suspending it from a new point $v_\star$, 
    then the resulting new triangulation is non-shellable and coincides with the patch around $v_\star$.
    This demonstrates that patches around boundary simplices are not necessarily shellable when the \cye{space} dimension \cye{$n$} is larger than three.
\end{remark}\todo{MV: Moved up a bit.}

@@

\begin{remark}
    Not all triangulable sets admit a triangulation that is also shellable. 
    Moreover, even if a set admits a shellable triangulation, not all of its triangulations might be shellable. 
    We refer to~\cite[Example 8.9]{ziegler2012lectures} for an examples of non-shellable triangulations of cubes and tetrahedra in three dimensions.

@@

\section{\cye{Outlook}} \todo{MV: Added this section at the end and moved here the corresponding contents from the introduction} \label{section:outlook}

There are several avenues for further research going \cye{further from the results established in this manuscript}.

\cye{Our} upper bounds are hardly the last word on estimating the Poincar\'e--Friedrichs constants over local patches (stars), where there are at least two points where the present techniques can be improved further. First, we utilize local Poincar\'e--Friedrichs constants over single simplices, subject to various boundary conditions, but these local estimates over convex \cye{\sout{domains}simplices} \todo{MV: ???} \cye{already} leave room for improvement.
Second, our construction needs to extend differential forms onto a simplex from the complement of that simplex inside a star. In the present manuscript, this is achieved via pullback along a bi-Lipschitz mapping. 
Better transformations or more careful estimates are likely possible, as are completely different approaches. For example, it is likely that further research on trace and extension operators on triangulations will enable better estimates.

Related to the first of the aforementioned opportunities for improvement, we conjecture that the Guerini--Savo estimates, hitherto only known for the exterior derivative as a mapping between Hilbert spaces and over convex domains with smooth boundary, could be extended to the entire range of Lebesgue exponents $1 \leq p \leq \infty$ and to convex Lipschitz domains. We believe such estimates would yield considerably better bounds on the Poincar\'e--Friedrichs constants than the regularized potential operators \cye{in the sprit of Costabel and McIntosh~\cite{costabel2010bogovskiui} that we employ now}. 
% We conjecture that the Poincar\'e--Friedrichs constant for the gradient operator still bounds the Poincar\'e--Friedrichs constants of the other instances of the exterior derivatives, as in \eqref{math:intro:guerinisavo}, when $1 \leq p \leq \infty$ and the domain is convex. 
On a related note, explicit estimates for the operator norms of regularized potential operators with different degrees of smoothing are of general interest.

\cye{Concerning the domains considered,} estimating Poincar\'e--Friedrichs constants over domains and manifolds with shellable triangulations is fundamentally restricted to \cye{\sout{spaces} domains} \todo{MV: ???} with simplicial homology groups of \disk s and spheres. In particular, domains with shellable triangulations are contractible. \cye{We believe that in the forthcoming works, t}hese estimates will \cye{become} subcomponent\cye{s} in computing upper bounds for Poincar\'e--Friedrichs constants over simplicial triangulations of general $n$-dimensional manifolds.

\mwl{
% Future work % DONE: Theo : parts already mentioned ??? 
% Computable reliable estimates are important: 
% they quantify the fundamental stability properties of partial differential equations in vector analysis,
% such as the Hodge Laplace equation. 
There are several avenues for further research going from here. 
The upper bounds in this manuscript are hardly the last word on estimating the Poincar\'e--Friedrichs constants over local patches. There are at least two points where the present techniques can be improved further. First, we utilize local Poincar\'e--Friedrichs constants over single simplices, subject to various boundary conditions, but these local estimates over convex domains likely leave room for improvement.
Second, our construction needs to extend differential forms onto a simplex from the complement of that simplex inside a local star. In the present manuscript, this is achieved via pullback along a bi-Lipschitz mapping. 
Better transformations or more careful estimates are likely possible, as are completely different approaches.  For example, it is likely that further research on trace and extension operators on triangulations will enable better estimates.

Estimating Poincar\'e--Friedrichs constants over domains and manifolds with shellable triangulations is fundamentally restricted to spaces with the simplicial homology groups of disks and spheres. In particular, domains with shellable triangulations are contractible. However, these estimates are important partial results. Forthcoming work will take these as a subcomponent in computing upper bounds for Poincar\'e--Friedrichs constants over simplicial triangulations of general $n$-dimensional manifolds. 
}
